---
title: "p8105_hw2_ew2717"
author: "Eric Wang"
output: github_document
---

```{r setup, echo = FALSE, message = FALSE}

library(tidyverse)
library(readxl)
library(lubridate)

```

## Problem 1

#### Importing data from Mr. Trash Wheel, cleaning the data, and then filtering it.

```{r importing, cleaning, filtering trash wheel}

trash_wheel_data <-
  read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
  sheet = "Mr. Trash Wheel",
  range = "A2:N408") %>% 
  janitor::clean_names() %>% 
  filter(!str_detect(month,'Total')) %>% 
  mutate(sports_balls = round(sports_balls))

```

#### Importing data from 2018 Precipitation, cleaning the data, and then filtering it.

```{r importing, cleaning, filtering 2018 precipitation}

precipitation_2018_data <-
  read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
  sheet = "2018 Precipitation",
  range = "A2:B15") %>% 
  janitor::clean_names() %>% 
  mutate(year = 2018) %>% 
  drop_na()

```

#### Importing data from 2019 Precipitation, cleaning the data, and then filtering it.

```{r importing, cleaning, filtering 2019 precipitation}

precipitation_2019_data <-
  read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
  sheet = "2019 Precipitation",
  range = "A2:B15") %>% 
  janitor::clean_names() %>% 
  mutate(year = 2019) %>% 
  drop_na()

```

#### Combining the data from 2018 Precipitation and 2019 Precipitation. 

```{r combining data}

precipitation_combined_data <-
  bind_rows(precipitation_2018_data, precipitation_2019_data) %>% 
  mutate(month = month.name[month]) %>% 
  relocate(year, month, total)

```

#### Data Observations.

```{r names total median mean rows and cols}

precipitation2018_names = precipitation_2018_data %>% names()
precipitation2019_names = precipitation_2019_data %>% names()
precipitation2018_total = precipitation_2018_data %>% summarise(sum(total))
precipitation2019_total = precipitation_2019_data %>% summarise(sum(total))
precipitation2018_median = precipitation_2018_data %>% summarise(median(total))
precipitation2019_median = precipitation_2019_data %>% summarise(median(total))
precipitation2018_mean = precipitation_2018_data %>% summarise(mean(total))
precipitation2019_mean = precipitation_2019_data %>% summarise(mean(total))
precipitation2018_row = precipitation_2018_data %>% nrow()
precipitation2019_row =  precipitation_2019_data %>% nrow()
precipitation2018_col = precipitation_2018_data %>% ncol()
precipitation2019_col = precipitation_2019_data %>% ncol()

```

The variable names for the 2018 precipitation data set are **`r precipitation2018_names`**, and the variable names for the 2019 precipitation data set are **`r precipitation2019_names`**. The total amount of precipitation in 2018 was **`r precipitation2018_total`**, and the total amount of precipitation in 2019 was **`r precipitation2019_total`**.The mean amount of precipitation in 2018 was **`r precipitation2018_mean`**, and the mean amount of precipitation in 2019 was **`r precipitation2019_mean`**. The median amount of precipitation in 2018 was **`r precipitation2018_median`**, and the median amount of precipitation in 2019 was **`r precipitation2019_median`**. The total number of rows in the 2018 precipitation data set is **`r precipitation2018_row`**, and the total number of rows in the 2019 precipitation data set is **`r precipitation2019_row`**. Lastly, the total number of columns in the 2018 precipitation data set is **`r precipitation2018_col`**, and the total number of columns in the 2019 precipitation data set is **`r precipitation2019_col`**.

## Problem 2

#### Importing data from FiveThirtyEight, specifically pols-month, and cleaning it.

```{r importing and cleaning pols-month}

pols_month_data <- 
  read_csv("data/fivethirtyeight_datasets/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(col = mon, into = c("year","month", "day")) %>% 
  mutate(across(.col = c(year, month, day), as.integer)) %>% 
  mutate(month = month.name[month]) %>% 
  mutate(president = case_when(prez_gop == 1 ~ "gop", prez_gop == 2 ~ "gop", prez_dem == 1 ~ "dem")) %>% 
  select(-day, -prez_dem, -prez_gop) %>% 
  select(year, month, president, everything())

```

#### Importing data from FiveThirtyEight, specifically snp, and cleaning it.

```{r importing and cleaning snp}

snp_data <- 
  read_csv("data/fivethirtyeight_datasets/snp.csv") %>%
  janitor::clean_names() %>% 
  separate(date, sep = "/", into = c("month", "day", "year")) %>%
  mutate(across(.cols = c("month", "day", "year"), as.integer)) %>%
  mutate(month = month.name[month]) %>%
  mutate(year = ifelse(year > 21, 1900 + year, 2000 + year)) %>%
  select(year, month, close)
  
```

#### Importing data from FiveThirtyEight, specifically unemployment, and cleaning it.

```{r importing and cleaning unemployment}

unemployment_data <- 
  read_csv("data/fivethirtyeight_datasets/unemployment.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
  jan:dec,
  names_to = "month", 
  values_to = "percentage") %>% 
  mutate(across(.col = c(year), as.integer)) %>% 
  mutate(month = recode(month, "jan" = "January", "feb" = "February", "mar" = "March", "apr" = "April", "may" = "May", "jun" = "June", "jul" = "July", "aug" = "August", "sep" = "September", "oct" = "October", "nov" = "November", "dec" = "December"))

```

#### Combining the pols and snp data.

```{r}

pols_snp_merged <-
  left_join(pols_month_data, snp_data, by = c("year", "month"))

```


#### Combining newly constructed data and umeployment data. 

```{r}

final_merged <-
  left_join(pols_snp_merged, unemployment_data, by = c("year", "month"))

```

#### Data Observations

```{r data observations for all three data sets}

pols_names = pols_month_data %>% names()
pols_dimension = pols_month_data %>%  dim()
pols_yrrange = pols_month_data %>% 
  select(year) %>% 
  range()

snp_names = snp_data %>% names()
snp_dimension = snp_data %>% dim()
snp_yrrange = snp_data %>% 
  select(year) %>% 
  range()

unemployment_names = unemployment_data %>% names()
unemployment_dimension = unemployment_data %>% dim()
unemployment_yrrange = unemployment_data %>% 
  select(year) %>% 
  range()

```

The pols-month data set contains data regarding politicians and whether they are democratic or republican. The important variables are **`r pols_names`**. The dimensions for the data set are **`r pols_dimension`**. The range of years for the data set are **`r pols_yrrange`**. 

The snp data set contains data that regarding Standard & Poor’s stock market index. The important variables are **`r snp_names`**. The dimensions for the data set are **`r snp_dimension`**. The range of years for the data set are **`r snp_yrrange`**. 

The unemployment data set contains data regarding the percentage of unemployment for specific years and months. The important variables are **`r unemployment_names`**. The dimensions for the data set are **`r unemployment_dimension`**. The range of years for the data set are **`r unemployment_yrrange`**. 


## Problem 3

#### Importing and cleaning baby name data. 

```{r importing then cleaning baby name data}

baby_name_data <-
  read_csv("data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  mutate(gender = str_to_sentence(gender),
  ethnicity = recode(ethnicity, 
                     "BLACK NON HISP" = "BLACK NON HISPANIC", 
                     "WHITE NON HISP" = "WHITE NON HISPANIC", 
                     "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER", 
                     "ASIAN AND PACIFIC" = "ASIAN AND PACIFIC ISLANDER"),
  ethnicity = str_to_title(ethnicity),
  childs_first_name = str_to_sentence(childs_first_name)
  )

```

#### Removing duplicate rows from baby name data.

```{r removing duplicate rows for baby name}

baby_name_data <-
    baby_name_data %>% distinct()

```

#### Producing well-structured and reader-friendly table showing the rank in popularity of the name “Olivia".

```{r producing table showing popularity of "olivia"}

olivia_table <-
  filter(baby_name_data, childs_first_name == "Olivia") %>%
  select(-count) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank") %>% 
  select(childs_first_name, gender, ethnicity, "2011", "2012", "2013", "2014", "2015", "2016")

knitr::kable(olivia_table)

```

#### Cleaning ranking table for male, white non-hispanic children born in 2016

```{r}

Male_rank <-
  filter(baby_name_data, rank == 1, gender == "Male") %>%
  select(-count) %>%
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "childs_first_name") %>%
  select(rank, gender, ethnicity, "2011", "2012", "2013", "2014", "2015", "2016")

knitr::kable(Male_rank)

```

#### Creating scatterplot Data 

```{r}

male_rank_data <-
  filter(baby_name_data, gender == "Male", ethnicity == "White Non Hispanic", year_of_birth == "2016") 
male_rank_data

view(male_rank_data)

```


###Creating Scatterplot

```{r}

male_rank_scatterplot = ggplot(male_rank_data, aes(x = rank, y = count)) + geom_point() + ggtitle("Name Popularity Rankings versus Count")

male_rank_scatterplot

```
