---
title: "p8105_hw2_ew2717"
author: "Eric Wang"
output: github_document
---

```{r setup, echo = FALSE, message = FALSE}

library(tidyverse)
library(readxl)
library(lubridate)

```

## Problem 1

#### Importing data from Mr. Trash Wheel, cleaning the data, and then filtering it.

```{r importing, cleaning, filtering trash wheel}

trash_wheel_data <-
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
  sheet = "Mr. Trash Wheel",
  range = "A2:N408") %>% 
  janitor::clean_names() %>% 
  filter(!str_detect(month,'Total')) %>% 
  mutate(sports_balls = round(sports_balls))

## view(trash_wheel_data)

```

#### Importing data from 2018 Precipitation, cleaning the data, and then filtering it.

```{r importing, cleaning, filtering 2018 precipitation}

precipitation_2018_data <-
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
  sheet = "2018 Precipitation",
  range = "A2:B15") %>% 
  janitor::clean_names() %>% 
  mutate(year = 2018) %>% 
  drop_na()

## view(precipitation_2018_data)

```

#### Importing data from 2019 Precipitation, cleaning the data, and then filtering it.

```{r importing, cleaning, filtering 2019 precipitation}

precipitation_2019_data <-
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
  sheet = "2019 Precipitation",
  range = "A2:B15") %>% 
  janitor::clean_names() %>% 
  mutate(year = 2019) %>% 
  drop_na()

## view(precipitation_2019_data)

```

#### Combining the data from 2018 Precipitation and 2019 Precipitation. 

```{r combining data}

precipitation_combined_data <-
  bind_rows(precipitation_2018_data, precipitation_2019_data) %>% 
  mutate(month = month.name[month]) %>% 
  relocate(year, month, total)

## view(precipitation_combined_data)

```

#### Finding the total precipitation for 2018 Precipitation and 2019 Precipitation.

```{r finding total precipitation}

total_precipitation_2018 = sum(precipitation_2018_data$total)
total_precipitation_2019 = sum(precipitation_2019_data$total)

```

Using the function above, we find that the total amount of precipitation in 2018 was `r total_precipitation_2018`.The total amount of precipitation in 2019 was `r total_precipitation_2019`.

#### Finding the median number of sports balls in a dumpster in 2019.

```{r finding median sports balls}



```

As seen above, the median number of sports balls in a dumpster in 2019 was r median_sportsballs_2019.

## Problem 2

